Product plots

### 2011-06-27 prodplots-infovis.tex
conditional and marginal density from joint

\begin{figure}[ht]
  \begin{tabular*}{0.5\linewidth}{rrr}
\multicolumn{3}{l}{$f(happiness, sex)$}\\
  \toprule
  & male & female \\ 
  \midrule
  not too happy & 0.05 & 0.07 \\ 
  pretty happy & 0.25 & 0.31 \\ 
  very happy & 0.14 & 0.18 \\ 
  \bottomrule
  \end{tabular*}
  \\[1em]

  \begin{tabular*}{0.48\linewidth}{rrr}
\multicolumn{3}{l}{$f(happiness| sex)$}\\
  \toprule
  & male & female \\
  \midrule
  not too happy & 0.43 & 0.57 \\
  pretty happy & 0.45 & 0.55 \\ 
  very happy & 0.43 & 0.57 \\
  \bottomrule
  \end{tabular*}
%
\hspace{1em}
  \begin{tabular*}{0.48\linewidth}{rrr}
\multicolumn{3}{l}{ $f(sex|happiness)$}\\
  \toprule
  & male & female \\ 
  \midrule
  not too happy & 0.12 & 0.12 \\ 
  pretty happy & 0.57 & 0.55 \\ 
  very happy & 0.31 & 0.32 \\
  \bottomrule
  \end{tabular*}
  \\[1em]

  \begin{tabular}{rrrr}
\multicolumn{3}{l}{$f(happiness), f(sex)$}\\
  \toprule
  & male & female & \\ 
  \midrule
  not too happy & & & 0.12 \\ 
  pretty happy & & & 0.56 \\ 
  very happy & & & 0.30 \\
  & 0.44 & 0.56 \\ 
  \bottomrule
  \end{tabular}

@@

Formally, a conditional distribution function is written $f(x | y)$ and is equal to $f(x, y) / f(y)$. This definition illustrates that an important statistical fact: given a conditional distribution and a marginal distribution, we can always find the joint distribution: $f(x, y) = f(x | y)\,f(y)$ \footnotemark. This definition extends in a straightforward way to higher dimensions. For example, a 3d joint distribution can be written as the product of 2d and 1d conditional and marginal distributions in the following three ways:

@@

This means that we can turn any high-dimensional {\sc pmf} into a product of low-dimensional conditional and marginal {\sc pmf}s, a computationally trivial operation. It remains to find a graphical analog of this multiplication.

>>>>

\begin{figure}[ht]
  \begin{tabular*}{0.5\linewidth}{rrr}
\multicolumn{3}{l}{$f(happy, sex)$, see fig. \ref{fig:fact-simple} (right) }\\
  \toprule
  & male & female \\ 
  \midrule
  not too happy & 0.05 & 0.07 \\ 
  pretty happy & 0.25 & 0.31 \\ 
  very happy & 0.14 & 0.18 \\ 
  \bottomrule
  \end{tabular*}
  \\[1em]

  \begin{tabular*}{0.48\linewidth}{rrr}
\multicolumn{3}{l}{$f(happy| sex)$}\\
  \toprule
  & male & female \\
  \midrule
  not too happy & 0.43 & 0.57 \\
  pretty happy & 0.45 & 0.55 \\ 
  very happy & 0.43 & 0.57 \\
  \bottomrule
  \end{tabular*}
%
\hspace{1em}
  \begin{tabular*}{0.48\linewidth}{rrr}
\multicolumn{3}{l}{ $f(sex|happy)$, see fig. \ref{fig:fact-simple} (middle) }\\
  \toprule
  & male & female \\ 
  \midrule
  not too happy & 0.12 & 0.12 \\ 
  pretty happy & 0.57 & 0.55 \\ 
  very happy & 0.31 & 0.32 \\
  \bottomrule
  \end{tabular*}
  \\[1em]

  \begin{tabular}{rrrr}
\multicolumn{3}{l}{$f(happy)$, see fig. \ref{fig:fact-simple} (left) and $f(sex)$}\\
  \toprule
  & male & female & \\ 
  \midrule
  not too happy & & & 0.12 \\ 
  pretty happy & & & 0.56 \\ 
  very happy & & & 0.30 \\
  & 0.44 & 0.56 \\ 
  \bottomrule
  \end{tabular}

@@

Formally, a conditional distribution function is written $f(x | y)$ and is equal to $f(x, y) / f(y)$. This definition illustrates that an important statistical fact: given a conditional distribution and a marginal distribution, we can always find the joint distribution: $f(x, y) = f(x | y)\,f(y)$\footnotemark. 
Conversely, we also see that we need to know both the marginal and conditional distributions in order to re-construct the joint distribution, i.e. only the joint distribution contains the full information about the relationship between all variables involved. Given the joint distribution $f(x,y)$, we can get either marginal distribution  by integrating (or summing, in the case of a categorical variable)  the other variable out: $f(x) = \int_y f(x,y) dy$. The marginal distribution together with the joint then lead to the conditional distribution. 
These definitions extend in a straightforward way to higher dimensions. For example, a 3d joint distribution can be written as the product of 2d and 1d conditional and marginal distributions in the following three ways:

@@

This means that we can build any high-dimensional {\sc pmf} as a product of low-dimensional conditional and marginal {\sc pmf}s, a computationally trivial operation. In the following, we are using the previously introduced low dimensional primitives  for a graphical analog of this multiplication.
